Data Exploration

### 2019-04-11 src/data_exploration.Rmd
minor alteration to make one-stop date change for running models

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
The "include=FALSE" supresses the package installation text appearing in the document...
```{r pre-install packages, include=FALSE}
packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap", "fastmatch", "tables") 
source("../src/ipak.R")
ipak(packages)
```

@@

and subset into background and presences
```{r}
presab <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
presence <- subset(presab, occurrence == "1")
background <- subset(presab, occurrence == "0")
```

@@

Temperature
```{r}
temp_depth_max <- max(presence$temp_depth, na.rm = TRUE)
temp_depth_min <- min(presence$temp_depth, na.rm = TRUE)
temp_depth_mean <- mean(presence$temp_depth, na.rm = TRUE)
temp_surface_max <- max(presence$temp_surface, na.rm = TRUE)
temp_surface_min <- min(presence$temp_surface, na.rm = TRUE)
temp_surface_mean <- mean(presence$temp_surface, na.rm = TRUE)
```

Salinity
```{r}
sal_depth_max <- max(presence$salinity_depth, na.rm = TRUE)
sal_depth_min <- min(presence$salinity_depth, na.rm = TRUE)
sal_depth_mean <- mean(presence$salinity_depth, na.rm = TRUE)
sal_surface_max <- max(presence$salinity_surface, na.rm = TRUE)
sal_surface_min <- min(presence$salinity_surface, na.rm = TRUE)
sal_surface_mean <- mean(presence$salinity_surface, na.rm = TRUE)
```

Chl
```{r}
chl_depth_max <- max(presence$chl_depth, na.rm = TRUE)
chl_depth_min <- min(presence$chl_depth, na.rm = TRUE)
chl_depth_mean <- mean(presence$chl_depth, na.rm = TRUE)
chl_surface_max <- max(presence$chl_surface, na.rm = TRUE)
chl_surface_min <- min(presence$chl_surface, na.rm = TRUE)
chl_surface_mean <- mean(presence$chl_surface, na.rm = TRUE)
```

O2
```{r}
o2_depth_max <- max(presence$o2_depth, na.rm = TRUE)
o2_depth_min <- min(presence$o2_depth, na.rm = TRUE)
o2_depth_mean <- mean(presence$o2_depth, na.rm = TRUE)
o2_surface_max <- max(presence$o2_surface, na.rm = TRUE)
o2_surface_min <- min(presence$o2_surface, na.rm = TRUE)
o2_surface_mean <- mean(presence$o2_surface, na.rm = TRUE)
```

@@

@@

# environmental correlates

@@

messing with code to see if can do for unique year month depth
```{r}
ascname <- list.files("../data/jan_test_86.9", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
noasc <- length(ascname)
  year <- "1998"
  month <- "01"
  depth <- "86.9294.asc"
  tempvar <- "temp"
  chlvar <- "chl"
  o2var <- "o2"
  salvar <- "salinity"
  depthvar <- "depth"
  surfacevar <- "surface"
  naovar <- "nao"
  amovar <- "amo"
  samplevar <- "sample"
  wintervar <- "winter"

}
```

# maxentvariableselection package

https://cran.r-project.org/web/packages/MaxentVariableSelection/MaxentVariableSelection.pdf

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
The "include=FALSE" supresses the package installation text appearing in the document...
```{r pre-install packages, include=FALSE}
packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap", "fastmatch", "tables", "dplyr") 
source("../src/ipak.R")
ipak(packages)
```

@@

and subset into background and presences
```{r}
presab <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
presence <- subset(presab, occurrence == "1")
background <- subset(presab, occurrence == "0")
```

@@

Temperature
```{r presence temp stats}
temp_depth_max <- max(presence$temp_depth, na.rm = TRUE)
temp_depth_min <- min(presence$temp_depth, na.rm = TRUE)
temp_depth_mean <- mean(presence$temp_depth, na.rm = TRUE)
temp_surface_max <- max(presence$temp_surface, na.rm = TRUE)
temp_surface_min <- min(presence$temp_surface, na.rm = TRUE)
temp_surface_mean <- mean(presence$temp_surface, na.rm = TRUE)
```

Salinity
```{r presence salinity stats}
sal_depth_max <- max(presence$salinity_depth, na.rm = TRUE)
sal_depth_min <- min(presence$salinity_depth, na.rm = TRUE)
sal_depth_mean <- mean(presence$salinity_depth, na.rm = TRUE)
sal_surface_max <- max(presence$salinity_surface, na.rm = TRUE)
sal_surface_min <- min(presence$salinity_surface, na.rm = TRUE)
sal_surface_mean <- mean(presence$salinity_surface, na.rm = TRUE)
```

Chl
```{r presence chl stats}
chl_depth_max <- max(presence$chl_depth, na.rm = TRUE)
chl_depth_min <- min(presence$chl_depth, na.rm = TRUE)
chl_depth_mean <- mean(presence$chl_depth, na.rm = TRUE)
chl_surface_max <- max(presence$chl_surface, na.rm = TRUE)
chl_surface_min <- min(presence$chl_surface, na.rm = TRUE)
chl_surface_mean <- mean(presence$chl_surface, na.rm = TRUE)
```

O2
```{r presence o2 stats}
o2_depth_max <- max(presence$o2_depth, na.rm = TRUE)
o2_depth_min <- min(presence$o2_depth, na.rm = TRUE)
o2_depth_mean <- mean(presence$o2_depth, na.rm = TRUE)
o2_surface_max <- max(presence$o2_surface, na.rm = TRUE)
o2_surface_min <- min(presence$o2_surface, na.rm = TRUE)
o2_surface_mean <- mean(presence$o2_surface, na.rm = TRUE)
```

@@

## plots of env. correlates by depth layer
the idea here is to see if there are any stark differences between the niche occupied by capelin at different depths

plot presences by unique depth & env. corr
```{r plot presences by unique depth & env. corr}
unique_depths <- unique(presence$depthlayer)
unique_depths

d1 <- subset(presence, depthlayerno == "1")
d10 <- subset(presence, depthlayerno == "10")
d11 <- subset(presence, depthlayerno == "11")
d12 <- subset(presence, depthlayerno == "12")
d13 <- subset(presence, depthlayerno == "13")
d14 <- subset(presence, depthlayerno == "14")
d15 <- subset(presence, depthlayerno == "15")
d16 <- subset(presence, depthlayerno == "16")
d17 <- subset(presence, depthlayerno == "17")
d18 <- subset(presence, depthlayerno == "18")
d19 <- subset(presence, depthlayerno == "19")
d20 <- subset(presence, depthlayerno == "20")
d21 <- subset(presence, depthlayerno == "21")
d22 <- subset(presence, depthlayerno == "22")
d23 <- subset(presence, depthlayerno == "23")
d24 <- subset(presence, depthlayerno == "24")
d25 <- subset(presence, depthlayerno == "25")
d26 <- subset(presence, depthlayerno == "26")
d27 <- subset(presence, depthlayerno == "27")
d28 <- subset(presence, depthlayerno == "28")
d29 <- subset(presence, depthlayerno == "29")
d30 <- subset(presence, depthlayerno == "30")
d31 <- subset(presence, depthlayerno == "31")
d32 <- subset(presence, depthlayerno == "32")
d33 <- subset(presence, depthlayerno == "33")
d34 <- subset(presence, depthlayerno == "34")
d35 <- subset(presence, depthlayerno == "35")
d36 <- subset(presence, depthlayerno == "36")
d37 <- subset(presence, depthlayerno == "37")
d38 <- subset(presence, depthlayerno == "38")
d39 <- subset(presence, depthlayerno == "39")
d40 <- subset(presence, depthlayerno == "40")
d41 <- subset(presence, depthlayerno == "41")
d42 <- subset(presence, depthlayerno == "42")
d43 <- subset(presence, depthlayerno == "43")
d44 <- subset(presence, depthlayerno == "44")
d45 <- subset(presence, depthlayerno == "45")
d46 <- subset(presence, depthlayerno == "46")
d47 <- subset(presence, depthlayerno == "47")

gtemp <- ggplot(d1, aes(x = temp_depth, colour = depth_layer)) + geom_density(na.rm = TRUE) + geom_density(data=d13, na.rm = TRUE) + geom_density(data=d14, na.rm = TRUE) + geom_density(data=d15, na.rm = TRUE) + geom_density(data=d16, na.rm = TRUE) + geom_density(data=d17, na.rm = TRUE) + geom_density(data=d18, na.rm = TRUE) + geom_density(data=d19, na.rm = TRUE) + geom_density(data=d20, na.rm = TRUE) + geom_density(data=d21, na.rm = TRUE) + geom_density(data=d22, na.rm = TRUE) + geom_density(data=d23, na.rm = TRUE) + geom_density(data=d24, na.rm = TRUE) + geom_density(data=d25, na.rm = TRUE) + geom_density(data=d26, na.rm = TRUE) + geom_density(data=d27, na.rm = TRUE) + geom_density(data=d28, na.rm = TRUE) + geom_density(data=d29, na.rm = TRUE) + geom_density(data=d30, na.rm = TRUE) + geom_density(data=d31, na.rm = TRUE) + geom_density(data=d32, na.rm = TRUE) + geom_density(data=d33, na.rm = TRUE) + geom_density(data=d34, na.rm = TRUE) + geom_density(data=d35, na.rm = TRUE) + geom_density(data=d36, na.rm = TRUE) + geom_density(data=d37, na.rm = TRUE) + geom_density(data=d38, na.rm = TRUE) + geom_density(data=d39, na.rm = TRUE) + geom_density(data=d40, na.rm = TRUE) + geom_density(data=d41, na.rm = TRUE) + geom_density(data=d42, na.rm = TRUE)
dev.copy(png,"../output/env/simple_plots/by_depth/temp_depth.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

#subtemp_depth <- subset(presence, select = c(depth_layer, temp_depth))

tdep <- plot(presence$temp_depth, presence$depth_layer)

ggplot(nafo0a, aes(x = temp_surface, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
dev.copy(png,"../output/env/env_by_nafo_plots/temp_surface_nafo.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

@@

@@
